---
title: "Using Kolada API with the lab5 Package"
author: "Christian Kammerer, Victor Guillo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Kolada API with the lab5 Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(here)
library(plotly)
```



# Introduction
The `lab5` package allows users to access data related to Swedish municipalities via the Kolada API. It includes functionality for data retrieval and visualization using R6 classes and integrates well with a Shiny application for interactive analysis. This vignette shows how to use the `KoladaHandler` and `MapHandler` classes in the package to fetch and visualize data.

## 1. KoladaHandler Class

The `KoladaHandler` class is responsible for communicating with the Kolada API and retrieving the relevant data.

### Creating a KoladaHandler Object
```{r}
library(lab5)
```


```{r}
# Initialize the KoladaHandler object
api_handler <- KoladaHandler$new()

# Check the available methods in KoladaHandler
print(api_handler)
```
```{r}
#Search for KPI
head(default_kpis)
```




```{r}
# Fetch KPI data for specific KPI IDs and municipalities
kpi_ids <- c("N00401")
municipality_ids <- c("0180", "0123") # Stockholm and another municipality
year <- 2020

kpi_result <- api_handler$get_data(kpi_ids, municipality_ids, year)
head(kpi_result)

```

```{r}
# Create a MapHandler instance
map_handler <- MapHandler$new()

```

```{r}
# Merge the KPI data with shapefile data
merged_data <- map_handler$merge_data(kpi_result)

# Plot the data on the map
map_handler$plot_data(merged_data, title = "KPI Map: Carbon Dioxide Emissions")

```



