---
title: "Using Kolada API with the lab5 Package"
author: "Christian Kammerer, Victor Guillo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Kolada API with the lab5 Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
library(here)
library(plotly)
```



# Introduction
The `lab5` package allows users to access data related to Swedish municipalities via the Kolada API. It includes functionality for data retrieval and visualization using R6 classes and integrates well with a Shiny application for interactive analysis. This vignette shows how to use the `KoladaHandler` and `MapHandler` classes in the package to fetch and visualize data.

## 1. KoladaHandler Class

The `KoladaHandler` class is responsible for communicating with the Kolada API and retrieving the relevant data.

### Creating a KoladaHandler Object
```{r}
library(lab5)

# Create a KoladaHandler instance
api_handler <- KoladaHandler$new()
```


```{r}
# Search for a KPI
kpi_data <- api_handler$parse_kpi("Carbon Dioxide Emissions per Capita")
head(kpi_data)

```


```{r}
# Fetch KPI data for specific KPI IDs and municipalities
kpi_ids <- c("N00401")
municipality_ids <- c("0180", "0123") # Stockholm and another municipality
year <- 2020

kpi_result <- api_handler$get_data(kpi_ids, municipality_ids, year)
head(kpi_result)

```

```{r}
# Create a MapHandler instance
map_handler <- MapHandler$new()

```

```{r}
# Merge the KPI data with shapefile data
merged_data <- map_handler$merge_data(kpi_result)

# Plot the data on the map
map_handler$plot_data(merged_data, title = "KPI Map: Carbon Dioxide Emissions")

```


```{r}
library(shiny)
library(lab5)

ui <- fluidPage(
  titlePanel("Municipality KPI Explorer"),
  sidebarLayout(
    sidebarPanel(
      selectInput("kpi", "Select KPI:", choices = c("Carbon Dioxide Emissions per Capita" = "N00401")),
      numericInput("year", "Select Year:", value = 2020, min = 2000, max = 2024),
      actionButton("update", "Update Map")
    ),
    mainPanel(
      plotlyOutput("kpi_map")
    )
  )
)

server <- function(input, output) {
  api_handler <- KoladaHandler$new()
  map_handler <- MapHandler$new()
  
  observeEvent(input$update, {
    kpi_data <- api_handler$get_data(input$kpi, municipality_ids = c("0180", "0123"), input$year)
    merged_data <- map_handler$merge_data(kpi_data)
    
    output$kpi_map <- renderPlotly({
      map_handler$plot_data(merged_data, title = "KPI Map: Carbon Dioxide Emissions")
    })
  })
}

shinyApp(ui, server)

```

